---
title: "ch01"
author: "B Davenport"
date: "29/04/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
```



```{r mpg}
mpg
?mpg

ggplot(data = mpg)+
    geom_point(mapping = aes(x = displ, y = hwy))
```
```{r}
# ggplot(data = <DATA>) +
#     <GEOM_FUNCTION>(mapping = aes(<MAPPING>))

ggplot(data = mpg)
# mpg has 234 rows and 11 columns/variables
# drv stands for drive train
ggplot(data = mpg)+
    geom_point(mapping = aes(hwy, cyl))
ggplot(data = mpg)+
    geom_point(mapping = aes(class, drv))
```
```{r}
ggplot(data = mpg) +
    geom_point(mapping = aes(x =displ, y =hwy, color = class))
```
## Excercises 3.3

```{r}
# ggplot(data = mpg) +
#     geom_point(mapping = aes(x =displ, y =hwy, size = class))
# ggplot(data = mpg) +
#     geom_point(mapping = aes(x =displ, y =hwy, alpha = class))
# ggplot(data = mpg) +
#     geom_point(mapping = aes(x =displ, y =hwy, shape = class))
# ggplot(data = mpg) +
#     geom_point(mapping = aes(x =displ, y =hwy), color = 'blue' )
# q.2
?mpg
glimpse(mpg)
# q.3
ggplot(data = mpg) +
    geom_point(mapping = aes(x =hwy, y = displ, color = cyl))
# a coninuous vairable cannot be mapped to shape.   
# ggplot(data = mpg) +
#     geom_point(mapping = aes(x =hwy, y = displ, shape = cyl))

ggplot(data = mpg) +
    geom_point(mapping = aes(x =hwy, y = displ, size = cyl))
# q.4
ggplot(data = mpg) +
    geom_point(mapping = aes(x =hwy, y =displ, size = cyl, color = cyl))
# q.5
ggplot(data = mpg) +
    geom_point(mapping = aes(x =hwy, y =displ, stroke = cyl))
# q.6
ggplot(data = mpg) +
    geom_point(mapping = aes(x =hwy, y =displ, color = displ < 5))
```
### 3.5 Facets
```{r}
ggplot(data = mpg) +
    geom_point(mapping = aes(x =displ, y =hwy)) +
    facet_wrap(drv~class, nrow =2)

ggplot(data = mpg) +
    geom_point(mapping = aes(x =displ, y =hwy)) +
    facet_wrap(.~class, nrow =2)
```
# 1. What happens when you facet a continuous variable
```{r}
ggplot(data = mpg) +
    geom_point(mapping = aes(x =drv, y =cyl)) +
    facet_grid(~displ)
```
2. What do the empty cells in plot with facet_grip(drv~cyl) have to do with the
other plot?
```{r}
ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy)) + 
  facet_grid(drv ~ cyl)

ggplot(data = mpg) +
    geom_point(mapping = aes(x =drv, y =cyl)) 
```

```{r}
ggplot(data = mpg) +
    geom_point(mapping = aes(x =displ, y =hwy)) +
    facet_grid(drv~ .)

ggplot(data = mpg) +
    geom_point(mapping = aes(x =displ, y =hwy)) +
    facet_grid(. ~ cyl)
```
facet_wrap allows us to see categories side by side when the number of thes cats
isn't too high. Usually you want the variable with the most variables in the columns
as this will make them more readable.
```{r}
ggplot(data = mpg) +
    geom_point(mapping = aes(x =displ, y =hwy)) +
    facet_wrap(~class, nrow = 2)
```
```{r}
ggplot(data = mpg) +
    geom_smooth(mapping = aes(x =displ, y =hwy, color = drv),
                show.legend = FALSE
    )
```

We can have multiple geoms within the same plot, it's as simple as 
geom()+
geom()
```{r}
ggplot(data = mpg) +
    geom_point(mapping = aes(x =displ, y =hwy)) +
    geom_smooth(mapping = aes(x = displ, y = hwy))
```
```{r}
ggplot(data = mpg, mapping = aes(x = displ, y = hwy)) +
    geom_point()+
    geom_smooth()
```
produces the same thing. if you place mappings in a geom function ggplot will overwrite
the global mappings for that layer only.
```{r}
ggplot(data = mpg, mapping = aes(x = displ, y = hwy)) +
    geom_point(mapping = aes(color = class)) +
    geom_smooth()
```
This can be used to specify different data for each layer here geom_smooth
is used to display the subcompact cars as a subset of the mpg data
```{r}
ggplot(data = mpg, mapping = aes(x = displ, y = hwy)) +
    geom_point(mapping = aes(color = class)) +
    geom_smooth(data = filter(mpg, class == "subcompact"), se = FALSE)
```
```{r}
ggplot(data = mpg, mapping = aes(x = displ, y = hwy, color = drv)) +
    geom_point(show.legend = FALSE)+
    geom_smooth(se = FALSE, show.legend = FALSE
    )
```


```{r 3.6}
ggplot(data = mpg,mapping = aes(x = displ, y = hwy)) +
  geom_point() +
  geom_smooth(se = FALSE)

ggplot(data = mpg,mapping = aes(x = displ, y = hwy)) +
  geom_point() +
  geom_smooth(mapping = aes(group = drv), se = FALSE)

ggplot(data = mpg,mapping = aes(x = displ, y = hwy, color = drv))+
  geom_point()+
  geom_smooth(se = FALSE)

ggplot(data = mpg,mapping = aes(x = displ, y = hwy)) +
  geom_point(mapping = aes(color = drv))+
  geom_smooth(se = FALSE)

ggplot(data = mpg,mapping = aes(x = displ, y = hwy))+
  geom_point(mapping = aes(color = drv))+
  geom_smooth(mapping = aes(linetype = drv), se = FALSE)

ggplot(data = mpg,mapping = aes(x = displ, y = hwy, color = drv))+
  geom_point()
```
# 3.7
Each geom has a stat count associated with it and visa versa, so you can usually just
use the geom with its associated stat by default, unless you want to edit the sta

```{r bars}
ggplot(data = diamonds)+
    geom_bar(mapping = aes(x = cut))
ggplot(data = diamonds)+
    stat_count(mapping = aes(x = cut))
```

change the stat

```{r demo}
demo <- tribble(
  ~cut,         ~freq,
  "Fair",       1610,
  "Good",       4906,
  "Very Good",  12082,
  "Premium",    13791,
  "Ideal",      21551
)

ggplot(data = demo) +
    geom_bar(mapping = aes(x = cut, y = freq), stat = "identity")
```
You can overwrite default mapping too
```{r}
ggplot(data = diamonds) +
    geom_bar(mapping = aes(x = cut, y = stat(prop), group = 1))
```
3.7 exercises
1.
```{r}
ggplot(data = diamonds)+
    stat_summary(
        mapping = aes(x = cut, y = depth),
        fun.min = min,
        fun.max = max,
        fun = median
    )
?stat_summary
# geom_pointrange
ggplot(data = diamonds) +
    geom_pointrange(mapping =aes(x = cut, y = depth),
      stat = "summary",
      fun.min = min,
      fun.max = max,
      fun = median
    )
```

2. geom_bar makes heights of the bars proportional to the number of cases, geom_col uses stat_identity() and leaves the data as it is.
```{r}
ggplot(data = diamonds)+
  geom_col(mapping = aes(x = cut, y = depth))
?geom_col
```

3. 

Geom     | stat
abline   | adds reference line
bar      | uses count
bin2d    | bin_2d
blank    | draws nothing used to ensure a common scale between plots
boxtour  | boxplot
contour  | contour visulaises a 3d surface in 2d
count    | count number of obs
density  | density smoothed histogram
density2d| same as above with filled contour bands density 2d
dotplot  | a bar plot with stacked dots or something like that
errorbar | errorbar identity
function | draws a function as a curve
hex      | binhex 
freqpoly | bin
jitter   | identity shortcut for geom_point(position = "jitter)
crossbar | identity representing a vertical interval
map      | identity pure annotation
path     | identity connects obs in the order they appear in data
point    | identity 
polygon  | identity like path but start and end connected and fill
quantile | quantile regression drawn with lines continuous boxplot
ribbon   | identity display interval by ymin + ymax
rug      | identity supplement  with marginal distribution
segment  | identity draws a straight line between points
smooth   | smooth aids the eye in seeing patterns
spoke    | identity useful for describing direction and distance
label    | identity adds a text label with a box
raster   | identity
violin   | ydensity a mirrored density plot in boxplot format

4. geom smooth computes
* y predicted value
* ymin lower pointwise CI around the mean
* ymax upper pointwise CI around the mean
* se standard error

5. Here we must use count as otherwise the bars are a proportion of themselves = 1 
```{r}
ggplot(data = diamonds) +
  geom_bar(mapping = aes(x = cut, y = after_stat(prop)))
ggplot(data = diamonds) +
  geom_bar(mapping = aes(x = cut, fill = color, y = after_stat(prop)))
ggplot(data = diamonds) +
  geom_bar(mapping = aes(x = cut, y = ..count.. / sum(..count..),
                         fill = color))
```

```{r}
ggplot(data = diamonds) +
  geom_bar(mapping = aes(x = cut, color = cut))
ggplot(data = diamonds) +
  geom_bar(mapping = aes(x = cut, fill = cut))
```
```{r}
ggplot(data = diamonds)+
  geom_bar(mapping = aes(x = cut, fill = clarity))
```

```{r}
ggplot(data = diamonds, mapping = aes(x = cut, fill = clarity)) +
  geom_bar(alpha = .2, position = "identity")
ggplot(data = diamonds, mapping = aes(x = cut, colour = clarity)) +
  geom_bar(fill = NA, position = "identity")
```
 position fill makes bars the same height this makes it easier to read proportions
```{r}
ggplot(data = diamonds,mapping = aes(x = cut, fill = clarity))+
  geom_bar(position = "fill")
```
positon dodge places overlapping objects beside one another
```{r}
ggplot(data = diamonds,mapping = aes(x = cut, fill = clarity)) +
  geom_bar(position = "dodge")


```
Jitter prevents overplotting allowing us to see where the mass of data lies.
```{r}
ggplot(data = mpg,mapping = aes(x = displ, y = hwy)) +
  geom_point(position = "jitter")
```
3.8 exercises
1. what is the problem with this plot how can it be improved?
```{r}
ggplot(data = mpg,mapping = aes(x = cty, y = hwy)) +
  # geom_point()
  geom_count(color = "blue") +
  geom_jitter()
?geom_jitter
```

```{r}
ggplot(data = mpg,mapping = aes(x = drv, y = hwy, color = class)) +
  geom_boxplot()
?geom_boxplot
```
coordinate systems are probably the most complicated part of ggplot2
default is cartesian
coor_flip() is useful for a horizontal boxplot or similar. This can be useful when
we have long labels.

```{r}
ggplot(data = mpg,mapping = aes(x = class, y = hwy))+
  geom_boxplot()
ggplot(data = mpg,mapping = aes(x = class, y = hwy)) +
  geom_boxplot() +
  coord_flip()
```
coord_quickmap() is flat coord_map() isn't
```{r}

nz <- map_data("nz")
ggplot(nz, aes(long, lat, group = group)) +
  geom_polygon(fill = "white", colour = "black") +
  coord_quickmap()
```
coor_polar() reveals an interesting connection between bar and coxcomb chart.
making a pie chart using theta 
```{r}
bar <- ggplot(data = diamonds) +
  geom_bar(
    mapping = aes(x = factor(1), fill = cut),
    show.legend = FALSE,
    width = 1) +
  theme(aspect.ratio = 1) +
  labs(x = NULL, y = NULL)

bar + coord_flip()
bar + coord_polar(theta = "y")

?geom_bar
```

ggplot(data = <DATA>) +
  <GEOM_FUNCTION>(
  mapping = aes(<MAPPING>),
  stat = <STAT>,
  position = <POSITION>
  ) +
  <COORDINATE_FUNCTION> +
  <FACET_FUNCTION>

The template above takes seven parameters, you rarely have to supply all seven as ggplot will usually supply a useful default.
These seven parameters compose the grammar of graphics
