---
title: "Monte Carlo"
author: "B Davenport"
date: "27/04/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# What is the Monte Carlo method
Repeatedly drawing samples from a known probability distribution, before calculating
the mean of these samples. Giving a good approximation of the final outcome.

```{r MC basic}
# Error in func1() : could not find function "func1"
# runs <- 100000
# func1 <- sum(sample(c(0,1), size = 10, replace = T)) >6
# mc_prob <- sum(replicate(runs, func1()))/runs
```

## Full Monte Carlo Sim in R 
* The winder can make 3000 – 5000 rolls per hour
* The bagger can make 2000 – 4000 rolls per hour
* The case packer can make 150 – 250 cases of 30 rolls each per hour
* The line will product at the slowest of the three
* Assume all distributions are uniform

```{r MC full}
results = NULL
for (k in 1:1000)
    {
     rolls = runif(1,3000,5000)
     bags = runif(1,2000,4000)
     cases = runif(1,150,200)*30
     total = min (rolls, bags, cases)
     results = rbind(results, data.frame(rolls, bags, cases, total))
    }
summary(results)
```

<https://web.stanford.edu/class/bios221/labs/simulation/lab_3_simulation.html>
This produces a differnt number each time as it is taking the mean of a 
distribution the lower the first number the greater the differences will be.
```{r}
vecpoisson=rpois(1,5)
mean(vecpoisson)
```
A sum of exponentials produces a gamma distribution, using an apply() function
apply() functions behave similarly to for loops but are much quicker way of itterating
over an object.
Here we generate 5 samples from an exponential rate of parameter 0.1 and sum them
together. sum(rexp(n=nexps, rate=rate)) replicate allows us to do this many times
with very little code 50000 here to give us an idea of the distribution.

```{r}
reps <- 50000
nexps <- 5
rate <- 0.1
set.seed(0)
system.time(
    x1 <- replicate(reps, sum(rexp(n=nexps, rate=rate)))
) #replicate
head(x1)
```
To confirm a gamma distribution lets make a histogram or qqplot of the simulation

```{r}
require(ggplot2)
ggplot(data.frame(x1), aes(x1)) +
    geom_histogram(aes(y=..density..)) +
    stat_function(fun=function(x)dgamma(x, shape = nexps, scale = 1/rate),
                 color = "red", size = 2)
```
```{r}
set.seed(0)
system.time(x1 <- lapply(1:reps, function(i){sum(rexp(n=nexps, rate = rate))})) # list apply
head(x1)
```

```{r}
set.seed(0)
system.time(x1 <- apply(matrix(rexp(n=nexps*reps, rate=rate), nrow=nexps),2,sum)) # apply on a matrix
head(x1)
```

```{r}
set.seed(0)
system.time(x1 <- colSums(matrix(rexp(n=nexps*reps, rate=rate), nrow=nexps))) # using colSums
```

```{r}
require(parallel)
set.seed(0)
system.time(x1 <- mclapply(1:reps, function(i){sum(rexp(n=nexps, rate=rate))})) # multi-cluster apply

```
## Predicting final standings of each premier league club.
<https://medium.com/swlh/predicting-the-outcome-of-the-english-premier-league-by-using-monte-carlo-method-in-r-638fbc90fd37>
```{r MC Football}
library(data.table)
library(stringr)
library(DT)

pl_fix <- fread("./epl-2020.csv")
datatable(pl_fix)

# some cleaning

# into new columns
pl_fix$result_home <- sub("-.*","", pl_fix$Result)
pl_fix$result_away <- sub(".*-","", pl_fix$Result)

convert_string_to_result <- function(x){
    
    x <- str_replace_all(string = x, pattern = " ", repl = "")
    x <- as.numeric(x)
    
    return(x)
    }

pl_fix$result_home <- convert_string_to_result(pl_fix$result_home)
pl_fix$result_away <- convert_string_to_result(pl_fix$result_away)

# create a draw d home h and away a indicator
pl_fix[, hda:= ifelse(result_home == result_away, "d",
                       ifelse(result_home > result_away, "h", "a"))]

team_names <- unique(c(pl_fix$`Home Team`, pl_fix$`Away Team`))
team_names

# creating a function that calculates how many points each team has in each
# round.

calculate_points <- function(pl_fix = pl_fix){
    
    # takes a table then calculates points for each round.
    round_table <- data.table(team = team_names, point = 0, matches_played = 0)
    
    for (row in c(1:nrow(pl_fix))){
        
        result <- pl_fix[row]$hda
        result <- as.character(result)
    }
}








```

